---
title: "Supplementary file 7. Global GO term similarity"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center"
)
options(width = 100)
```

```{r}
library(simplifyEnrichment)
se_opt$verbose = FALSE
```

In this supplementary, we look at the global GO term similarities for Biological Process (BP) ontology.
First we calculate the similarity matrix for all GO BP terms supported in _GOSemSim_ package.

```{r, eval = !file.exists("all_bp_sim.rds")}
library(GOSemSim)
semData = godata("org.Hs.eg.db", ont = "BP")
ic = semData@IC

library(simplifyEnrichment)
# This takes very long time
sm = GO_similarity(names(ic), measure = "Rel", ont = "BP")
```


```{r, eval = !file.exists("all_bp_sim.rds"), echo = FALSE}
saveRDS(sm, file = "all_bp_sim.rds")
```

```{r, echo = FALSE, eval = file.exists("all_bp_sim.rds")}
sm = readRDS("all_bp_sim.rds")
```

The number of all GO terms in BP ontology:

```{r}
dim(sm)
```

## GO BP terms can be clustered into several major groups

To simplify the similarity matrix, the GO terms which have zero similarity to all other terms are removed.

```{r}
sm_tmp = sm
diag(sm_tmp) = 1
l = rowSums(sm_tmp > 0) > 1
sm = sm[l, l]
rm(sm_tmp)
```

Next we construct the GO tree. The parent-children information is from _GO.db_ package. The
`GOBPCHILDREN` data object contains lists of direct children terms for every GO terms. If a GO term
is a leaf in the GO tree, the value for it is `NA`.

```{r}
library(GO.db)
lt = as.list(GOBPCHILDREN)
lt[1:5]
```

Then we can construct a graph object for the GO tree.

```{r}
df = do.call("rbind", lapply(names(lt), function(nm) {
	data.frame(rep(nm, length(lt[[nm]])), lt[[nm]])
}))
df2 = df[!is.na(df[, 2]), ]

library(igraph)
g = graph.edgelist(as.matrix(df2), directed = TRUE)
g
```

We take the common GO terms in the GO tree and the global similarity matrix.

```{r}
cn = intersect(rownames(sm), unlist(df2))
sm2 = sm[cn, cn]
g = induced_subgraph(g, cn)
sm2 = sm2[V(g)$name, V(g)$name]
```

The final number of GO terms are:

```{r}
dim(sm2)
```

The gobal similarity matrix is clustered by the "louvain" method. Note binary cut is too slow
to work on a similarity matrix with 15k GO terms.

```{r, eval = !file.exists("all_bp_sim_cl.rds")}
# This takes around 30 min
cl = cluster_terms(sm2, method = "louvain")
```

```{r, echo = FALSE, eval = !file.exists("all_bp_sim_cl.rds")}
saveRDS(cl, file = "all_bp_sim_cl.rds")
```

```{r, echo = FALSE, eval = file.exists("all_bp_sim_cl.rds")}
cl = readRDS("all_bp_sim_cl.rds")
```

The size of clusters:

```{r}
table(cl)
```

To visualize the similarity matrix and the clustering, we only randomly sample 2000 terms.

```{r, fig.width = 8, fig.height = 4}
set.seed(123)
ind = sample(length(cl), 2000)
ht_clusters(sm2[ind, ind], cl = cl[ind], exclude_words = c("process", "regulation", "response"))
```

## GO clusters are dominated by several top GO terms

Next we look at what are the top GO terms in each cluster (_i.e._, the GO terms located
on the top of the GO subtree in each cluster). First we construct subgraph for each cluster.

```{r}
gl = tapply(rownames(sm2), cl, function(go_id) {
    induced_subgraph(g, go_id)
})
```

Since the subgraphs are split from the global tree/graph, they might have several disconnected compoments.
We check the sizes of disconnected components in each subgraph.

```{r}
lapply(gl, function(g) {
    table(components(g)$membership)
})
```

As we can see, in each cluster, there is always a huge components which have majority of the terms while
other components have very few terms. Thus, we only take the componets with the largest number of terms.

```{r}
gl2 = lapply(gl, function(g) {
    membership = components(g)$membership
    induced_subgraph(g, membership == which.max(table(membership)))
})
```

Next we extract the top GO terms where the in-degree is zero. And we only want the GO terms
which are the parent nodes of a huge GO subtree. The influences a parent node gives to all its 
descendant nodes is measured by the ["spread centrality"](https://rdrr.io/github/jokergoo/CePa/man/spread.html).

```{r}
library(CePa)
lapply(gl2, function(g) {
    s1 = degree(g, mode = "in")
    s2 = spread(g, mode = "out")
    s2[s1 == 0]
})
```

As shown above, in each cluster, there are some top GO terms that are only the
parents of a small number of children terms while some other terms have extremely 
high spread values which means they are parents of huge number of children terms.
We simply remove the top GO terms with spread values less than the mean spread values
of all top GO terms.

```{r, echo = FALSE}
options(width = 40)
```

```{r}
go_list = lapply(gl2, function(g) {
    s1 = degree(g, mode = "in")
    s2 = spread(g, mode = "out")
    go_id = V(g)$name[which(s1 == 0)]
    s = s2[s1 == 0]
    go_id = go_id[s >= mean(s)]
    term = suppressMessages(select(GO.db::GO.db, keys = go_id, columns = "TERM")$TERM)
    paste0(go_id, ": ", term)
})
```

The final major top GO terms in each cluster are:

```{r}
go_list
```
