---
title: "Supplementary file 3. Examples of semantic similarity measures"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center",
    echo = FALSE
)
```

In this supplementary, we demonstrate the heatmaps of semantic similarity matrices from
different semantic measures, which are "Rel", "Resnik", "Lin", "Jiang" and "Wang" (see 
[the GOSemSim vignette](https://bioconductor.org/packages/release/bioc/vignettes/GOSemSim/inst/doc/GOSemSim.html) for more details of the methods). The matrices are calculated based on 500 randomly sampled GO terms, from Biological Process (BP), Molecular Function (MF) and Cellular Component (CC) ontologies. For each ontology category, we show 5 examples.

Generally, the similarity matrices generated by the "Rel", "Resnik" and "Lin" methods show
more clear diagonal block patterns. 

```{r}
library(simplifyEnrichment)
library(grid)
library(circlize)
library(ComplexHeatmap)
library(GetoptLong)
library(cowplot)

set.seed(123)
```

## 500 random BP {.tabset}

### Random 1

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500)
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
library(cowplot)
plot_grid(plotlist = pl, nrow = 1)
```


### Random 2

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500)
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 3

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500)
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 4

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500)
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 5

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500)
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```




## 500 random MF {.tabset}

### Random 1

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "MF")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
library(cowplot)
plot_grid(plotlist = pl, nrow = 1)
```


### Random 2

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "MF")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 3

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "MF")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 4

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "MF")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 5

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "MF")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```



## 500 random CC {.tabset}

### Random 1

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "CC")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
library(cowplot)
plot_grid(plotlist = pl, nrow = 1)
```


### Random 2

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "CC")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 3

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "CC")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 4

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "CC")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```


### Random 5

```{r, fig.width = 10*1.5, fig.height = 10/5*1.2*1.5}
go_id = random_GO(500, "CC")
pl = list()
for(measure in c("Rel", "Resnik", "Lin", "Jiang", "Wang")) {
	mat = GO_similarity(go_id, measure = measure)
	col_fun = colorRamp2(range(mat), c("white", "red"))
	pl[[measure]] = grid.grabExpr(draw(Heatmap(mat, col = col_fun, show_row_names = FALSE, 
		name = qq("@{measure} similarity"), show_column_names = FALSE,
		show_row_dend = FALSE, show_column_dend = FALSE, 
		column_title = qq("measure = @{measure}"),
		heatmap_legend_param = list(direction = "horizontal", title_position = "lefttop", legend_width = unit(3, "cm"))),
		heatmap_legend_side = "bottom"))
}
plot_grid(plotlist = pl, nrow = 1)
```

